On discretely entropy conservative discontinuous Galerkin methods

### 2017-07-29 docs/splitform/main.tex
committing updates to entropy stable DG paper

        % Draw slope triangle.
        \draw[#5]   (A)-- node[pos=0.5,anchor=north] {1}
                    (B)-- 
                    (C)-- node[pos=0.5,anchor=west] {#4}
                    cycle;
    }
}
%%% END MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.

@@

        % Draw slope triangle.
        \draw[#5]   (A)-- node[pos=.5,anchor=south] {1}
                    (B)-- 
                    (C)-- node[pos=0.5,anchor=west] {#4}
                    cycle;
    }
}
%%% END MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.

@@

        % Draw slope triangle.
        \draw[#5]   (A)-- node[pos=0.5,anchor=east] {#4}
                    (B)-- 
                    (C)-- node[pos=0.5,anchor=south] {1}
                    cycle;
    }
}
%%% END MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.

@@

The accuracy observed for high order methods can be attributed in part to their low numerical dissipation and dispersion properties \cite{ainsworth2004dispersive}.  While low numerical dissipation is advantageous for problems in linear wave propagation \cite{komatitsch1998spectral}, the high numerical dissipation present in low order schemes adds a stabilizing effect for nonlinear hyperbolic problems \cite{wang2013high}.  As a result, high order schemes for nonlinear conservation laws tend to require additional stabilization, including (but not limited to) filtering \cite{hesthaven2007nodal}, slope limiting \cite{krivodonova2007limiters}, artificial viscosity \cite{persson2006sub}, and polynomial de-aliasing \cite{ karniadakis2013spectral, kirby2003aliasing}.  Moreover, stabilized numerical methods can still fail, requiring user intervention or heuristic modifications to achieve non-divergent solutions.

@@

The construction of entropy stable high order schemes was done much more recently for the compressible Euler and Navier-Stokes equations in \cite{fisher2013high, carpenter2014entropy} based on a spectral collocation approach on tensor product elements, which can be interpreted as a mass-lumped DG spectral element (DG-SEM) scheme.  The proof of entropy stability relies on the presence of a diagonal mass matrix, as well as the summation by parts property \cite{gassner2013skew}.  Similar entropy-stable schemes were constructed for the shallow water and MHD equations \cite{derigs2016novel, gassner2016well, wintermeyer2017entropy}.  More recently, entropy stable schemes were extended to unstructured triangular meshes in \cite{chen2017entropy} using the split-form methodology of \cite{gassner2016split} and multi-dimensional SBP operators \cite{hicken2016multidimensional}.

@@

This work focuses on the construction of entropy conservative high order DG schemes for systems of conservation laws.  In order to generalize beyond collocation DG and diagonal norm SBP operators, we will consider dense (non-lumped) mass matrices and quadrature rules of arbitrarily high accuracy (over-integration).  The outline of the paper is as follows: Section~\ref{sec:intro} will briefly review the construction of entropy conservative collocation methods on a single element.  Section~\ref{sec:ecdg} will describe how to construct analogous entropy conservative methods on single element.  Section~\ref{sec:ecdg2} will discuss extensions to multiple elements, including comparisons of different coupling terms and entropy stable fluxes.  We note that, for simplicity and clarity of presentation, the proposed methods and proofs are given in one dimension.  However, the extension to multiple dimensions (including simplicial elements) is straightforward and will be described in an upcoming manuscript.  
Finally, Section~\ref{sec:num} presents numerical results which verify the high order accuracy and discrete entropy conservation of the proposed methods.

@@

We will begin by reviewing continuous entropy theory, as well as existing high order methods which are provably entropy conservative at the discrete level.  These methods were introduced in \cite{fisher2013high, carpenter2014entropy} for tensor product elements, though we adopt methods of proof introduced more recently in \cite{gassner2017br1, chen2017entropy}.

%\note{Specify what we mean by diagonal norm collocation DG method.}

@@

Using that $\bm{B}$ is diagonal, combining the latter term in (\ref{eq:sbpentropy}) with boundary terms involving the numerical flux yields
\[
\sum_{i,j = 1}^{N+1} \bm{B}_{ij}\bm{v}_i^T \bm{f}_S(\bm{u}_i,\bm{u}_j) + \bm{v}_i^T\bm{B}_{ij}\LRp{\bm{f}_i^* - \bm{f}(\bm{u}_i)} = \bm{1}^T\bm{B} \LRp{\bm{v}^T \bm{f}^*},
\]
where we have used the consistency of $\bm{f}_S(\bm{u},\bm{u}) = \bm{f}(\bm{u})$.  Combining everything results in a discrete statement of the conservation of entropy (\ref{eq:consentropy}).   
%Combining everything results in a discrete statement of the conservation of entropy (\ref{eq:consentropy}) 
%\[
%\bm{1}^T\LRp{\bm{M}\pd{U(\bm{u})}{t} + \bm{B}\LRp{\bm{v}^T\bm{f}^* - \bm{\psi}}} = 0.
%\]
\end{proof}
We note that, unlike the proof of continuous entropy conservation, the proof of discrete entropy conservation does not require the use of the chain rule, which may not hold at the discrete level.

The proof of entropy conservation can also be extended beyond a single element by specifying appropriate simultaneous approximation terms (SBP-SAT).  This can be done by taking $\bm{f}_i^*$ to be the entropy conservative flux $f_S(u_L,u_R)$ at the interface between two elements \cite{carpenter2014entropy, gassner2017br1, chen2017entropy}.  Similarly, employing an entropy stable flux at element interfaces results in an entropy stable method which satisfies a global entropy inequality.  
%\end{remark}

@@

We assume that the domain $\Omega \in \mathbb{R}^d$ is exactly triangulated by a mesh which consists of the union of non-overlapping elements $D^k$.  We further assume that each element $D^k$ is the image of a reference element $\widehat{D}$ under the local elemental mapping 
\[
x^k = \bm{\Phi}^k \widehat{x},
\]
where $x^k$ denote physical coordinates on $D^k$ and $\widehat{x}$ denote coordinates on the reference element.  Finally, we identify $J$ as the determinant of the Jacobian of $\bm{\Phi}^k$, and refer to it as the Jacobian for the remainder of this work.  For clarity and simplicity of presentation, we assume $\bm{\Phi}^k$ is affine in this case, such that $J$ is constant over each element $D^k$.

@@

It is helpful to first understand flux differencing as a method for recovering split formulations.  We illustrate this using the one-dimensional Burgers equation on domain $ [-1,1]$, which is given in conservative form as 
\[
\pd{u}{t} + \pd{f(u)}{x} = 0, \qquad f(u) = \frac{1}{2}u^2.
\]
%An alternative form of the Burgers' equation can be derived by taking a convex combination of the conservative and non-conservative forms of the spatial term
%\begin{equation}
%\pd{u}{t} + \frac{1}{3}\LRp{\pd{u^2}{x} + u\pd{u}{x}} = 0.
%\label{eq:burgerssplit}
%\end{equation}
Assuming smooth $u$, continuous energy conservation can be shown by multiplying by $u$, integrating over $[-1,1]$, and using $\pd{u}{t}u = \frac{1}{2}\pd{u^2}{t}$
\begin{equation}
\frac{1}{2}\pd{}{t}\int_{-1}^1{u}^2 = -\left.\frac{u^3}{3}\right|_{-1}^1.
\label{eq:burgersconservation}
\end{equation}

@@

Direct computations also show that $\LRp{D_N 1,vw}_{\widehat{D}} = 0$ for all bounded weights $w(x)$ and $\forall v \in V_h$.  Thus, we can apply (\ref{eq:fluxdiff}) with $D_N$ to construct a discrete formulation for Burgers' equation
\begin{align}
\LRp{\pd{u}{t},v} + \frac{1}{3}\LRp{D_N u^2 + uD_Nu,v} &= \LRp{\pd{u}{t},v}_{\widehat{D}} + \frac{1}{3}\LRp{D_N u^2,v}_{\widehat{D}} + \frac{1}{3}\LRp{D_Nu,uv}_{\widehat{D}},
%\LRp{\pd{u}{t},v}_{\widehat{D}} + \frac{1}{3}\LRp{\pd{\Pi_N u^2}{x},v}_{\widehat{D}} + \frac{1}{3}\LRp{D_Nu,uv}_{\widehat{D}}
\qquad \forall v\in V_h.
\label{eq:burgersdiscrete}
\end{align}
To show this formulation is discretely energy conservative, we take $v=u$ and use Lemma~\ref{lemma:dgd_local} to cancel volume terms, yielding
\[
\frac{1}{2}\pd{}{t}\LRp{u,u}_{\widehat{D}} = -\frac{1}{3}\LRa{u^2,u}_{\partial \widehat{D}},
\]
which is a discrete analogue of the conservation condition (\ref{eq:burgersconservation}).  This is equivalent to the skew-symmetric split formulation described in \cite{gassner2013skew}; however, as noted in \cite{chen2017entropy}, it can be derived from flux differencing without explicit knowledge of the skew form.  We will show later in Section~\ref{sec:recovery} that a natural extension of this formulation to multiple elements recovers the weak form of the method introduced in \cite{ranocha2017extended}.

@@

For the proof, we will need the following assumption:
\begin{assumption}
The numerical flux $\bm{f}_S(x,y)$ admits an absolutely convergent expansion of the form
\[
\bm{f}_S(x,y) = \sum_{i=1}^{\infty} \bm{f}_i(x) \bm{g}_i(y), 
\]
where $\bm{f}_i, \bm{g}_i$ are differentiable.  
\label{eq:assumption}
\end{assumption}
This assumption is satisfied, for example, by $\bm{f}_S$ which are analytic and admit a convergent Taylor expansion in $x$ and $y$.  In Appendix~\ref{appendix:A}, we show that this assumption also holds for the choices of flux functions $\bm{f}_S$ considered in this work, assuming the solution satisfies physical bound constraints.  However, we do need this assumption in computation implementations of the proposed methods.

We now introduce a discretization for the nonlinear system of conservation laws (\ref{eq:pde}) on $[-1,1]$: find $\bm{u}\in V_h$ such that
\begin{equation}
\LRp{\pd{\bm{u}}{t},\bm{w}}_{\widehat{D}} + \LRp{\left.\LRp{2D_N \bm{f}_S(\bm{u}_x,\bm{u}_y)}\right|_{y=x},\bm{w}}_{\widehat{D}}= \LRa{\bm{f}^*-\bm{f}(\bm{u}_x),\bm{w}}_{\partial \widehat{D}}, \qquad \bm{w}\in V_h,
\label{eq:ecdg_oneelem}
\end{equation}
where $\bm{f}^*$ is a numerical flux used to impose boundary conditions weakly, and $\bm{u}_x,\bm{u}_y$ are evaluations of $\bm{u}$ in terms of the projections of entropy variables $\bm{v}(\bm{u}(x))$ 
\begin{equation}
%\bm{u}_x = \bm{u}(\Pi_N\bm{v}(\bm{u}(x))), \qquad \bm{u}_y = \bm{u}(\Pi_N\bm{v}(\bm{u}(y))).
\bm{u}_x(x) = \bm{u}\LRp{(\Pi_N\bm{v})(x)}, \qquad \bm{u}_y(y) = \bm{u}\LRp{(\Pi_N\bm{v})(y)}.
\label{eq:projectedentropyvars}
\end{equation}

@@

&\LRp{\left.\LRp{D_N \sum_{i=1}^{\infty}\bm{f}_i(\bm{u}_x)\bm{g}_i(\bm{u}_y)}\right|_{y=x}, \Pi_N\bm{v}(x)}_{\widehat{D}} = \sum_{i=1}^{\infty} \LRp{{D_N \bm{f}_i(\bm{u}_x)}, {\bm{g}_i(\bm{u}_x)\Pi_N\bm{v} }}_{\widehat{D}}\\
&= \sum_{i=1}^{\infty}\LRa{\bm{f}_i(\bm{u}_x) \bm{g}_i(\bm{u}_x),\Pi_N\bm{v}}_{\partial \widehat{D}}  - \LRp{ \bm{f}_i(\bm{u}_x), D_N \LRp{\bm{g}_i\Pi_N\bm{v} }}_{\widehat{D}}\\
&= \LRa{\bm{f}_S(\bm{u}_x,\bm{u}_x), \Pi_N\bm{v}(x)}_{\partial \widehat{D}}  - \LRp{\left.\LRp{D_N \Pi_N\LRp{\bm{f}_S(\bm{u}_x,\bm{u}_y)\bm{v}(x)} }\right|_{y=x},1}_{\widehat{D}}\\
&= \LRa{\bm{f}(\bm{u}_x), \Pi_N\bm{v}(x)}_{\partial \widehat{D}}  - \LRp{\left.\LRp{D_N \Pi_N\LRp{\bm{f}_S(\bm{u}_x,\bm{u}_y)\bm{v}(x)} }\right|_{y=x},1}_{\widehat{D}},

@@

The volume terms can be manipulated to yield

@@

&\LRp{\left.\LRp{D_N \LRp{\bm{f}_S(\bm{u}_x,\bm{u}_y)\LRp{\Pi_N\bm{v}(y)-\Pi_N\bm{v}(x)} }}\right|_{y=x},1}_{\widehat{D}} \\

@@

We see that, in the proof of Theorem~\ref{lemma:ec1}, recovering the time derivative of entropy on the left hand side requires testing with the projection of the entropy variables.  For collocation-type DG methods, $L^2$ projection and interpolation are equivalent, which implies that evaluating the flux $\bm{f}_S$ using point-wise nodal values of the solution satisfies the entropy conservation condition given in Definition~\ref{def:tadmor}.  For collocation-type DG methods, the method can also be simplified by noting that $\bm{V}_q,\bm{P}_q$ are identity matrices.

In general, the projection of the entropy variables does not agree with their point-wise evaluation.  However, the entropy conservation condition given in Definition~\ref{def:tadmor} holds so long as the flux is evaluated in a manner consistent with the entropy variables.  This motivates (\ref{eq:projectedentropyvars}) as a non-standard way of evaluating the flux, which is the primary difference between the construction of entropy conservative schemes for collocation-type DG methods and projection-based DG methods.

@@

\note{It's currently unclear how to distinguish whether different operators are high order accurate. }

\note{Kopriva's note ``Stability of Overintegration Methods for Nodal Discontinuous Galerkin Spectral Element Methods'' does not apply here, because the formulation is entropy stable for any quadratures which evaluate $P^N$ exactly (assuming affine meshes!).  Certain terms are under-integrated, but they do not contribute to the interplay between volume and surface terms described by Kopriva.}

\subsection{Conservation of entropy on multiple elements}

Extending the entropy stable DG formulation (\ref{eq:ecdg_oneelem}) is done by combining the flux differencing with the global DG differential operator $D^x_h$
\begin{equation}
\sum_{k}\LRp{\pd{\bm{u}}{t},\bm{w}}_{D^k} + \LRp{\left.\LRp{2D^x_h\bm{f}_S(\bm{u}_x,\bm{u}_y)}\right|_{y=x},\bm{w}}_{D^k} = \LRa{\bm{f}(\bm{u}_x)-\bm{f}^*,\bm{w}}_{\partial \Omega},
%\int_{D^k}\pd{\bm{u}}{t} \bm{w} + \left.\LRp{2D^x_h\bm{f}_S\LRp{\bm{u}(\Pi_N\bm{v}({x})),\bm{u}(\Pi_N\bm{v}({y}))}}\right|_{{y}={x}}\bm{w} + \LRa{\bm{f}^*-\bm{f}(\bm{u}),\bm{w}}_{\partial \Omega}= 0, \qquad \forall \bm{w}\in V_h^n,
\label{eq:ecdg}
\end{equation}
where $\bm{f}^*$ is some numerical flux used in the imposition of boundary conditions, and the global operator $D^x_h$ is understood to be applied to each component of $\bm{f}_S\LRp{\bm{u}(x),\bm{u}(y)}$.

\note{Add assumptions about $\bm{u}(x)$ - needs to be locally analytic.}

%\note{Global DG operators reveal how to construct SBP-SAT terms for any collocation-type DG methods (where $L^2$ projection and interpolation are equivalent).  Not easy to deal with SBP-SAT terms for more complex systems \cite{ranocha2017extended, ranocha2017comparison}.  }
This formulation  yields a theorem analogous to Theorem~\ref{lemma:ec1}:
\begin{theorem}
The formulation (\ref{eq:ecdg}) is globally entropy conservative in the sense that
\[
\LRp{\pd{U(\bm{u})}{t},1}_{\Omega} =  \LRa{\psi,1}_{\partial \Omega}-\LRa{{\bm{f}^*, \Pi_N\bm{v}}}_{\partial \Omega}.
\]
\end{theorem}
\begin{proof}
The proof is identical to the proof of Theorem~\ref{lemma:ec1}, except that Lemma~\ref{lemma:ibp} is used in place of Lemma~\ref{lemma:dgd_local}.
\end{proof}

@@

\appendix
\section{Expansion of the flux function}
\label{appendix:A}

>>>>

        % Draw slope triangle.
        \draw[#5]   (A)-- node[pos=0.5,anchor=north] {}
                    (B)-- 
                    (C)-- node[pos=0.5,anchor=west] {#4}
                    cycle;
    }
}
%%% END MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.

@@

        % Draw slope triangle.
        \draw[#5]   (A)-- node[pos=.5,anchor=south] {}
                    (B)-- 
                    (C)-- node[pos=0.5,anchor=west] {#4}
                    cycle;
    }
}
%%% END MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.

@@

        % Draw slope triangle.
        \draw[#5]   (A)-- node[pos=0.5,anchor=east] {#4}
                    (B)-- 
                    (C)-- node[pos=0.5,anchor=south] {}
                    cycle;
    }
}
%%% END MACRO FOR ANNOTATION OF TRIANGLE WITH SLOPE %%%.

@@

The accuracy observed for high order methods can be attributed in part to their low numerical dissipation and dispersion properties \cite{ainsworth2004dispersive}.  While low numerical dissipation is advantageous for problems in linear wave propagation \cite{chan2015gpu}, the high numerical dissipation present in low order schemes adds a stabilizing effect for nonlinear hyperbolic problems \cite{wang2013high}.  As a result, high order schemes for nonlinear conservation laws tend to require additional stabilization, including (but not limited to) filtering \cite{hesthaven2007nodal}, slope limiting \cite{krivodonova2007limiters}, artificial viscosity \cite{persson2006sub}, and polynomial de-aliasing \cite{ karniadakis2013spectral, kirby2003aliasing}.  Moreover, stabilized numerical methods can still fail, requiring user intervention or heuristic modifications to achieve non-divergent solutions.

@@

The construction of entropy conservative high order schemes was done much more recently for the compressible Euler and Navier-Stokes equations in \cite{fisher2013high, carpenter2014entropy} based on a spectral collocation approach on tensor product elements, which can be interpreted as a mass-lumped DG spectral element (DG-SEM) scheme.  The proof of entropy conservation relies on the presence of a diagonal mass matrix, as well as the summation by parts property \cite{gassner2013skew}.  Similar entropy-stable schemes were constructed for the shallow water and MHD equations \cite{gassner2016well, wintermeyer2017entropy,winters2017uniquely}.  More recently, entropy conservative and entropy stable schemes were extended to unstructured triangular meshes in \cite{chen2017entropy} using the split-form methodology of \cite{gassner2016split} and multi-dimensional SBP operators \cite{hicken2016multidimensional}.

@@

This work focuses on the construction of entropy conservative high order DG schemes for systems of conservation laws.  In order to generalize beyond collocation DG and diagonal norm SBP operators, we will consider dense (non-lumped) mass matrices and more general quadrature rules (e.g.\ rules without boundary points, rules with more points than the dimension of the corresponding approximation space).  We will present proofs of discrete entropy stability in terms of integrals and $L^2$ projections, which we assume are computed using an appropriate quadrature rule.  We stress that, while we utilize a continuous notation involving integrals rather than discrete sums, the proofs rely only on properties of quadrature-based integration and $L^2$ projection.

The outline of the paper is as follows: Section~\ref{sec:intro} will briefly review the construction of entropy conservative collocation methods on a single element.  Section~\ref{sec:ecdg} will describe how to construct analogous entropy conservative methods on single element.  Section~\ref{sec:ecdg2} will discuss extensions to multiple elements, including comparisons of different coupling terms and entropy stable fluxes.  We note that, for simplicity and clarity of presentation, the proposed methods and proofs are given in one dimension.  However, the extension to multiple dimensions (including simplicial elements) is straightforward and will be described in an upcoming manuscript.  
Finally, Section~\ref{sec:num} presents numerical results which verify the high order accuracy and discrete entropy conservation of the proposed methods.

@@

We will begin by reviewing continuous entropy theory, as well as existing high order methods which are provably entropy conservative at the discrete level.  These methods were introduced in \cite{fisher2013high, carpenter2014entropy} for tensor product elements, though we adopt methods of proof introduced more recently in \cite{gassner2017br1, chen2017entropy}.  These methods all share a basis in diagonal norm summation by parts operators.  Because they can be derived from a nodal discontinuous Galerkin framework through the collocation of nodal points and quadrature points, we refer to them as ``diagonal norm collocation DG methods'' in this work.

@@

Using that $\bm{B}$ is diagonal, combining the latter term in (\ref{eq:sbpentropy}) with boundary terms involving the numerical flux yields
\[
\sum_{i,j = 1}^{N+1} \bm{B}_{ij}\bm{v}_i^T \bm{f}_S(\bm{u}_i,\bm{u}_j) + \bm{v}_i^T\bm{B}_{ij}\LRp{\bm{f}_i^* - \bm{f}(\bm{u}_i)} = \bm{1}^T\bm{B} \LRp{\bm{v}^T \bm{f}^*},
\]
where we have used the consistency of $\bm{f}_S(\bm{u},\bm{u}) = \bm{f}(\bm{u})$.  Combining everything results in a discrete statement of the conservation of entropy (\ref{eq:consentropy}).   
%Combining everything results in a discrete statement of the conservation of entropy (\ref{eq:consentropy}) 
%\[
%\bm{1}^T\LRp{\bm{M}\pd{U(\bm{u})}{t} + \bm{B}\LRp{\bm{v}^T\bm{f}^* - \bm{\psi}}} = 0.
%\]
\end{proof}
We note that, unlike the proof of continuous entropy conservation, the proof of discrete entropy conservation does not require the use of the chain rule, which may not hold at the discrete level.  This proof of entropy conservation can also be extended beyond a single element by specifying appropriate simultaneous approximation terms (SBP-SAT).  This can be done by taking $\bm{f}_i^*$ to be the entropy conservative flux $f_S(u_L,u_R)$ at the interface between two elements \cite{carpenter2014entropy, gassner2017br1, chen2017entropy}.  Similarly, employing an entropy stable flux at element interfaces results in an entropy stable method which satisfies a global entropy inequality.

It is important to emphasize that conservation of entropy does not, in general, guarantee stability of the numerical scheme unless the solution satisfies additional constraints.  For example, the convex entropy associated with the compressible Euler and Navier-Stokes equations is well-defined only if the density and pressure are positive, and steps must be taken in any numerical scheme to guarantee that the discrete solution of such constraints.  For DG methods, this is most commonly done using bound and positivity-preserving limiters, such as those introduced in \cite{zhang2010positivity, zhang2012maximum}.

@@

\note{Fix notation: switch $V_h$ to $P^N$}
We assume that the domain $\Omega \in \mathbb{R}^d$ is exactly triangulated by a mesh which consists of the union of non-overlapping elements $D^k$.  We further assume that each element $D^k$ is the image of a reference element $\widehat{D}$ under the local elemental mapping 
\[
x^k = \bm{\Phi}^k \widehat{x},
\]
where $x^k$ denote physical coordinates on $D^k$ and $\widehat{x}$ denote coordinates on the reference element.  Finally, we identify $J$ as the determinant of the Jacobian of $\bm{\Phi}^k$, and refer to it as the Jacobian for the remainder of this work.  For clarity and simplicity of presentation, we assume $\bm{\Phi}^k$ is affine in this case, such that $J$ is constant over each element $D^k$.

@@

It is helpful to first understand flux differencing as a method for recovering split formulations.  We illustrate this using the one-dimensional Burgers equation on domain $ [-1,1]$, which is given in conservative form as 
\[
\pd{u}{t} + \pd{f(u)}{x} = 0, \qquad f(u) = \frac{1}{2}u^2.
\]
%An alternative form of the Burgers' equation can be derived by taking a convex combination of the conservative and non-conservative forms of the spatial term
%\begin{equation}
%\pd{u}{t} + \frac{1}{3}\LRp{\pd{u^2}{x} + u\pd{u}{x}} = 0.
%\label{eq:burgerssplit}
%\end{equation}
Assuming smooth $u$, Burgers' equation conserves a square entropy $U(\bm{u}) = u^2/2$.   This can be shown by multiplying by $u$, integrating over $[-1,1]$, and using $\pd{u}{t}u = \frac{1}{2}\pd{u^2}{t}$
\begin{equation}
\frac{1}{2}\pd{}{t}\int_{-1}^1{u}^2 = -\left.\frac{u^3}{3}\right|_{-1}^1.
\label{eq:burgersconservation}
\end{equation}

@@

Direct computations also show that $\LRp{D_N 1,vw}_{\widehat{D}} = 0$ for all bounded weights $w(x)$ and $\forall v \in V_h$.  Thus, we can apply (\ref{eq:fluxdiff}) with $D_N$ to construct a discrete formulation for Burgers' equation
\begin{align}
\LRp{\pd{u}{t},v} + \frac{1}{3}\LRp{D_N u^2 + uD_Nu,v} &= \LRp{\pd{u}{t},v}_{\widehat{D}} + \frac{1}{3}\LRp{D_N u^2,v}_{\widehat{D}} + \frac{1}{3}\LRp{D_Nu,uv}_{\widehat{D}},
%\LRp{\pd{u}{t},v}_{\widehat{D}} + \frac{1}{3}\LRp{\pd{\Pi_N u^2}{x},v}_{\widehat{D}} + \frac{1}{3}\LRp{D_Nu,uv}_{\widehat{D}}
\qquad \forall v\in V_h.
\label{eq:burgersdiscrete}
\end{align}
To show this formulation is discretely energy conservative, we take $v=u$ and use Lemma~\ref{lemma:dgd_local} to cancel volume terms, yielding
\[
\frac{1}{2}\pd{}{t}\LRp{u,u}_{\widehat{D}} = -\frac{1}{3}\LRa{u^2,u}_{\partial \widehat{D}},
\]
which is a discrete analogue of the conservation condition (\ref{eq:burgersconservation}).  This is equivalent to the skew-symmetric split formulation described in \cite{gassner2013skew}; however, as noted in \cite{chen2017entropy}, it can be derived from flux differencing without explicit knowledge of the skew form.  We will show later in Section~\ref{sec:recovery} that a natural extension of this formulation to multiple elements recovers the weak form of the scheme introduced in \cite{ranocha2017extended}.

@@

For the proof, we will need the following assumption:
\begin{assumption}
The numerical flux function $\bm{f}_S$ admits an absolutely convergent expansion of the form 
\[
\bm{f}_S(x,y) = \sum_{i=1}^{\infty} \bm{f}_i(x) \bm{g}_i(y), 
\]
where $\bm{f}_i, \bm{g}_i$ are real and bounded over each element $D^k$.  
\label{eq:assumption}
\end{assumption}
This assumption is satisfied, for example, by $\bm{f}_S$ which are real analytic and thus admit a Taylor expansion in $x$ and $y$.  In Appendix~\ref{appendix:A}, we show that this assumption also holds for the choices of flux functions $\bm{f}_S$ considered in this work, assuming the solution satisfies physical bound constraints.  However, we note that this assumption is not required in the computation implementations of the proposed methods, and may be an overly restrictive condition on $\bm{f}_S$.

We now introduce a discretization for the nonlinear system of conservation laws (\ref{eq:pde}) on $[-1,1]$: find $\bm{u}\in V_h$ such that
\begin{equation}
\LRp{\pd{\bm{u}}{t},\bm{w}}_{\widehat{D}} + \LRp{\left.\LRp{2D_N \bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y))}\right|_{y=x},\bm{w}}_{\widehat{D}}= \LRa{\bm{f}^*-\bm{f}(\bm{u}_{\bm{v}}),\bm{w}}_{\partial \widehat{D}}, \qquad \bm{w}\in V_h,
\label{eq:ecdg_oneelem}
\end{equation}
where $\bm{f}^*$ is a numerical flux used to impose boundary conditions weakly, and $\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y)$ are evaluations of $\bm{u}$ in terms of the projections of entropy variables $\bm{v}(\bm{u}(x))$ 
\begin{equation}
%\bm{u}_{\bm{v}}(x) = \bm{u}(\Pi_N\bm{v}(\bm{u}(x))), \qquad \bm{u}_{\bm{v}}(y) = \bm{u}(\Pi_N\bm{v}(\bm{u}(y))).
\bm{u}_{\bm{v}}(x) = \bm{u}\LRp{(\Pi_N\bm{v})(x)}.
\label{eq:projectedentropyvars}
\end{equation}

@@

&\LRp{\left.\LRp{D_N \sum_{i=1}^{\infty}\bm{f}_i(\bm{u}_{\bm{v}}(x))\bm{g}_i(\bm{u}_{\bm{v}}(y))}\right|_{y=x}, \Pi_N\bm{v}(x)}_{\widehat{D}} = \sum_{i=1}^{\infty} \LRp{{D_N \bm{f}_i(\bm{u}_{\bm{v}})}, {\bm{g}_i(\bm{u}_{\bm{v}})\Pi_N\bm{v} }}_{\widehat{D}}\\
&= \sum_{i=1}^{\infty}\LRa{\bm{f}_i(\bm{u}_{\bm{v}}) \bm{g}_i(\bm{u}_{\bm{v}}),\Pi_N\bm{v}}_{\partial \widehat{D}}  - \LRp{ \bm{f}_i(\bm{u}_{\bm{v}}), D_N \LRp{\bm{g}_i\Pi_N\bm{v} }}_{\widehat{D}}\\
&= \LRa{\bm{f}_S(\bm{u}_{\bm{v}},\bm{u}_{\bm{v}}), \Pi_N\bm{v}(x)}_{\partial \widehat{D}}  - \LRp{\left.\LRp{D_N \LRp{\bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y))\Pi_N\bm{v}(x)} }\right|_{y=x},1}_{\widehat{D}}\\
&= \LRa{\bm{f}(\bm{u}_{\bm{v}}), \Pi_N\bm{v}(x)}_{\partial \widehat{D}}  - \LRp{\left.\LRp{D_N\LRp{\bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y))\Pi_N\bm{v}(x)} }\right|_{y=x},1}_{\widehat{D}},

@@

Noting that 
\left.\LRp{D_N \bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y))}\right|_{y=x}\bm{v}(x) = \left.\LRp{D_N \LRp{\bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y))\bm{v}(y)}}\right|_{y=x},
the volume terms can be manipulated to yield

@@

&\LRp{\left.\LRp{D_N \LRp{\bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y))\LRp{\Pi_N\bm{v}(y)-\Pi_N\bm{v}(x)} }}\right|_{y=x},1}_{\widehat{D}} \\

@@

We see that, in the proof of Theorem~\ref{lemma:ec1}, recovering the time derivative of entropy on the left hand side requires testing with the projection of the entropy variables.  For collocation-type DG methods, $L^2$ projection and interpolation are equivalent, which implies that evaluating the flux $\bm{f}_S$ using point-wise nodal values of the solution satisfies the entropy conservation condition given in Definition~\ref{def:tadmor}.  For diagonal norm collocation-type DG methods, the method can also be simplified by noting that the projection operator reduces to a nodal interpolation operator.

In general, the projection of the entropy variables does not agree with their point-wise evaluation.  However, the entropy conservation condition given in Definition~\ref{def:tadmor} holds so long as the flux is evaluated in a manner consistent with the entropy variables.  This motivates (\ref{eq:projectedentropyvars}) as a non-standard way of evaluating the flux, which is the primary difference between the construction of entropy conservative schemes for diagonal norm collocation-type DG methods and projection-based DG methods.

\begin{remark}
In the proof of Theorem~\ref{lemma:ec1}, Assumption~\ref{eq:assumption} is used only to formally apply Lemma~\ref{lemma:dgd_local}.  The use of this assumption can be avoided by proving instead a discrete conservation of entropy in terms of a fixed quadrature rule and the quadrature-based matrices introduced in Section~\ref{sec:implementation}.  
\end{remark}

@@

%\note{It's currently unclear how to distinguish whether different operators are high order accurate. }
%\note{Kopriva's note ``Stability of Overintegration Methods for Nodal Discontinuous Galerkin Spectral Element Methods'' does not apply here, because the formulation is entropy stable for any quadratures which evaluate $P^N$ exactly (assuming affine meshes!).  Certain terms are under-integrated, but they do not contribute to the interplay between volume and surface terms described by Kopriva.}

\subsection{Conservation of entropy and entropy stability on multiple elements}
\label{sec:multipleelems}
Extending the entropy stable DG formulation (\ref{eq:ecdg_oneelem}) is done by combining the flux differencing with the global DG differential operator $D^x_h$
\begin{equation}
\sum_{k}\LRp{\pd{\bm{u}}{t},\bm{w}}_{D^k} + \LRp{\left.\LRp{2D^x_h\bm{f}_S(\bm{u}_{\bm{v}}(x),\bm{u}_{\bm{v}}(y))}\right|_{y=x},\bm{w}}_{D^k} = \LRa{\bm{f}(\bm{u}_{\bm{v}})-\bm{f}^*,\bm{w}}_{\partial \Omega},
%\int_{D^k}\pd{\bm{u}}{t} \bm{w} + \left.\LRp{2D^x_h\bm{f}_S\LRp{\bm{u}(\Pi_N\bm{v}({x})),\bm{u}(\Pi_N\bm{v}({y}))}}\right|_{{y}={x}}\bm{w} + \LRa{\bm{f}^*-\bm{f}(\bm{u}),\bm{w}}_{\partial \Omega}= 0, \qquad \forall \bm{w}\in V_h^n,
\label{eq:ecdg}
\end{equation}
where $\bm{f}^*$ is some numerical flux used in the imposition of boundary conditions, and the global operator $D^x_h$ is understood to be applied to each component of $\bm{f}_S\LRp{\bm{u}(x),\bm{u}(y)}$.

%\note{Add assumptions about $\bm{u}(x)$ - needs to be locally analytic.}

%\note{Global DG operators reveal how to construct SBP-SAT terms for any collocation-type DG methods (where $L^2$ projection and interpolation are equivalent).  Not easy to deal with SBP-SAT terms for more complex systems \cite{ranocha2017extended, ranocha2017comparison}.  }
This formulation yields a theorem analogous to Theorem~\ref{lemma:ec1}:
\begin{theorem}
\label{thm:ecdg_global}
The formulation (\ref{eq:ecdg}) is globally entropy conservative in the sense that
\[
\LRp{\pd{U(\bm{u})}{t},1}_{\Omega} =  \LRa{\psi,1}_{\partial \Omega}-\LRa{{\bm{f}^*, \Pi_N\bm{v}}}_{\partial \Omega}.
\]
\end{theorem}
\begin{proof}
The proof is identical to the proof of Theorem~\ref{lemma:ec1}, except that Lemma~\ref{lemma:ibp} is used in place of Lemma~\ref{lemma:dgd_local}.
\end{proof}

@@

\appendix
\section{Assumptions on the flux function}
\label{appendix:A}
